---
title: "Week 2020/49: Toronto Shelters"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
extrafont::loadfonts(device="win")
#extrafont::fonttable()
#extrafont::font_import("C:/Windows/Fonts/", pattern = "RobotoCondensed")

library(tidyverse)
library(lubridate)
library(ggtext)
```

```{r}
raw_df <- tidytuesdayR::tt_load(2020, week = 49)
raw_df <- raw_df$shelters
```
```{r}
raw_df
```

```{r}
raw_df %>%
  distinct(sector)
```

There appears to be different shelters based on need. 

Let's take a look at occupancy with respect to sector in  raw numbers:

```{r}
raw_df %>%
  mutate(month = month(occupancy_date, label = TRUE),
         year = year(occupancy_date)) %>%
  group_by(month, year, sector) %>%
  summarise(mean_occupancy = mean(occupancy, na.rm = T)) %>%
  ggplot(aes(x = month, y = mean_occupancy, colour = fct_reorder(sector, mean_occupancy, .desc = TRUE), group = sector)) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year) +
  theme_minimal() + 
   theme(
     axis.text.x = element_text(family = "Roboto Condensed", size = 11, colour = "grey30", angle = 90, vjust = 0.5),
     axis.text.y = element_text(family = "Roboto Condensed", size = 11, colour = "grey30"),
     plot.title = element_text(family = "Roboto Condensed", size = 18, colour = "#000000", face = "bold"),
     plot.subtitle = element_text(family = "Roboto Condensed", size = 16, colour = "#000000"),
     panel.grid.minor = element_blank(),
     panel.grid.major = element_line(linetype = "dotted"),
     strip.text.x = element_textbox(
      size = 12,
      color = "white", fill = "#5D729D", box.color = "#4A618C",
      halign = 0.5, linetype = 1, r = unit(1, "pt"), width = unit(1, "npc"),
      padding = margin(2, 0, 1, 0), margin = margin(3, 3, 3, 3)),
     legend.title = element_blank(),
     
   ) +
  labs(
    title = "Toronto Shelter Occupacy",
    subtitle = "Occupancy for families dramatically increased from 2017 into 2018",
    legend = "",
    x = "",
    y = "Monthly Mean Occupancy"
  ) +
  colorspace::scale_color_discrete_qualitative()
```

What about as occupancy as a fraction of capacity?

```{r}
raw_df %>%
  mutate(month = month(occupancy_date, label = TRUE),
         year = year(occupancy_date)) %>%
  group_by(month, year, sector) %>%
  summarise(mean_occupancy = mean(occupancy, na.rm = T),
            mean_capacity = mean(capacity, na.rm = T),
            pct = mean_occupancy / mean_capacity) %>%
ggplot(aes(x = month, y = pct*100, colour = as.factor(sector), group = sector)) +
  geom_line() +
  geom_point() +
  facet_grid(. ~ year) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  ) +
  scale_x_discrete(
    name = "Month"
  ) +
  scale_y_continuous(
    name = "Percent occupied (%)",
    limits = c(70, 100)
  )
```

And is there any spare capacity in the network?

```{r dpi = 300, fig.height = 6, fig.width = 8}
raw_df %>%
  mutate(diff = capacity - occupancy) %>%
   mutate(month = month(occupancy_date, label = TRUE),
         year = year(occupancy_date)) %>%
  group_by(month, year, sector) %>%
  summarise(mean_diff = mean(diff, na.rm = T), .groups = "keep") %>%
  ungroup() %>%
  ggplot(aes(x = month, y = mean_diff)) +
  geom_col(fill = "#5D729D", colour = NA) +
  facet_grid(sector ~ year) +
  labs(
    title = "Spare Capacity in Toronto Shelters",
    subtitle = "What is the difference between capacity and occupancy?",
    x = "**Month**",
    y = "**Spare capacity**"
  ) +
  theme_minimal() + 
   theme(
     aspect.ratio = 0.3,
     axis.text.x = element_text(family = "Roboto Condensed", size = 11, colour = "grey30", angle = 90, vjust = 0.5),
     axis.text.y = element_text(family = "Roboto Condensed", size = 11, colour = "grey30"),
     axis.title.y = element_markdown(family = "Roboto Condensed", colour = "grey30", margin = margin(0, 3, 0, 0, "pt"), hjust = 1, vjust = 1),
     axis.title.x = element_markdown(family = "Roboto Condensed", colour = "grey30", margin = margin(3, 0, 0, 0, "pt"), hjust = 1, vjust = 1),
     plot.title = element_text(family = "Roboto Condensed", size = 18, colour = "#000000", face = "bold"),
     plot.subtitle = element_text(family = "Roboto Condensed", size = 16, colour = "#000000"),
     panel.grid.major = element_line(size = 0.25),
     panel.grid.minor = element_blank(),
     strip.text.x = element_textbox(
      size = 12,
      color = "white", fill = "#5D729D", box.color = "#4A618C",
      halign = 0.5, linetype = 1, r = unit(1, "pt"), width = unit(1, "npc"),
      padding = margin(2, 0, 1, 0), margin = margin(3, 3, 3, 3)),
     strip.text.y = element_blank()
   ) +
  scale_y_continuous(
    expand = c(0, 0.1),
    breaks = seq(0, 20, 10)
  ) +
  geom_richtext(
    aes(x = 0.5,
        y = 20,
        hjust = 0,
        label = if_else((year == 2017 & month == "Jan"), glue::glue("**{sector}**"), NA_character_)),
    label.colour = NA,
    family = "Roboto Condensed", 
    size = 4, 
    colour = "grey30"
  )
```

Seems like a lot of spare capcity for families but very little for people who fall in the other categories. Given occupancy is near 100%, this suggests the demand is there for shelters that cater to individuals. Perhaps resources should be diverted to these other categories? Is monthly occupancy remarkably consistent for individuals because there is no change in demand or simply no available shelters? 